\documentclass[sigconf, review, anonymous, 9pt]{acmart}

\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{tabularx}
\usepackage{makecell}
\usepackage{pifont}
\usepackage{threeparttable}
\usepackage[referable]{threeparttablex}
\setlength{\marginparwidth}{2cm}

\usepackage[textsize=footnotesize]{todonotes}
\presetkeys{todonotes}{inline,backgroundcolor=yellow}{}

\renewcommand\theadfont{\bfseries}
\newcolumntype{L}{>{\raggedright\arraybackslash}X}
\newcolumntype{C}{>{\centering\arraybackslash}X}
\usepackage{xspace}
\usepackage{xcolor}
\usepackage{longtable}
\hypersetup{colorlinks=true,linkcolor=black,urlcolor=blue,citecolor=black}

\newcommand{\parag}[1]{\vspace{2mm}\noindent\textbf{#1.}\hspace{2mm}}

% Draft scaffolding: mark stale, pending-rewrite content in blue. Remove for camera-ready.
\newcommand{\outdatedbanner}{\par\noindent\textcolor{blue}{\textbf{[OUTDATED]}}\par\medskip}


% Metrics
\newcommand{\SCI}{\texttt{SCI}\xspace}
\newcommand{\RSCI}{\texttt{rSCI}\xspace}
\newcommand{\OSCI}{\texttt{oSCI}\xspace}
\newcommand{\TSCI}{\texttt{tSCI}\xspace}

\newcommand{\E}{\texttt{E}\xspace}
\newcommand{\I}{\texttt{I}\xspace}
\newcommand{\M}{\texttt{M}\xspace}
\newcommand{\R}{\texttt{R}\xspace}


% Appendix status taxonomy colours
\newcommand{\statusA}[1]{\textcolor[HTML]{1a7f37}{\textbf{#1}}}
\newcommand{\statusB}[1]{\textcolor[HTML]{9a6700}{#1}}
\newcommand{\statusC}[1]{\textcolor[HTML]{d1242f}{#1}}


% Coverage symbols: \yes (full), \maybe (partial / w/ caveat), \no (excluded), \na (not applicable)
\newcommand{\yes}{\ding{51}}
\newcommand{\no}{\ding{55}}
\newcommand{\maybe}{\ensuremath{\circ}}
\newcommand{\na}{\textendash}

% Lever-response marks for the sustainability-levers table (\S\ref{sec:sota}):
% \faithful (moves with real carbon), \proxyonly (moves, wrong driver), \notrefl (no response),
% \perverse (improves although real carbon does not fall).
\newcommand{\faithful}{\textcolor[HTML]{1a7f37}{\ding{51}}}
\newcommand{\proxyonly}{\textcolor[HTML]{9a6700}{\ensuremath{\sim}}}
\newcommand{\notrefl}{\textcolor[HTML]{d1242f}{\ding{55}}}
\newcommand{\perverse}{\textcolor[HTML]{8250df}{\ensuremath{\downarrow}}}

% Takeaway box (no box package loaded otherwise): boxed, italic, for the \S2 closer.
\newsavebox{\takeawaybox}
\newenvironment{takeaway}
  {\par\medskip\noindent\begin{lrbox}{\takeawaybox}%
   \begin{minipage}{\dimexpr\columnwidth-2\fboxsep-2\fboxrule\relax}%
   \small\textbf{Takeaway.}\hspace{1mm}\itshape}
  {\end{minipage}\end{lrbox}\noindent\fbox{\usebox{\takeawaybox}}\par\medskip}

\AtBeginDocument{%
  \providecommand\BibTeX{{Bib\TeX}}}

\setcopyright{none}

\begin{document}

%\title{Accounting for Action: Reconciling Per-Workload Carbon Metrics with Corporate GHG Reporting}
\title{Reconciling Bottom-Up Metrics with Top-Down Reporting for Carbon Accounting in the Cloud}

\author{Philipp Wiesner}
\affiliation{%
  \institution{TBD}
  \country{TBD}
}
\email{TBD}

\begin{abstract}
There exist two largely separate approaches to quantifying the carbon footprint of cloud workloads.
Bottom-up metrics such as Software Carbon Intensity (\SCI) provide real-time signals for energy-efficient and carbon-aware optimization, but are blind to provider-side overheads.
Top-down corporate carbon reports and customer-facing dashboards provide a comprehensive view of the tenant's carbon footprint from cloud usage, but are coarse, delayed, and methodologically opaque.
Because the two approaches differ in scope and reporting frequency, the current state of the art treats them as decoupled: bottom-up metrics for optimization, top-down numbers for corporate disclosure.

We argue that although the two signals can never be identical, \emph{they should be reconcilable}. 
If organizations cannot trace their bottom-up improvements through to the audited top-down report, they will lose trust in both metrics. 
In this paper, we survey the state of the art in cloud carbon accounting and propose a vision for \RSCI: a metric that re-anchors a bottom-up energy estimate to the cloud provider's per-scope, location-based report through a residual bridge, which can be allocated to workloads in a way that incentivizes sustainable behavior.
% \RSCI aligns the two signals by construction: Scheduling metrics contain provider-side overheads, and degrades gracefully when a provider does not yet ship reconciliation-grade data.
%The residual can itself be allocated to workloads in ways that incentivize sustainable behavior---for example, penalizing demand peaks that force new hardware procurement---which no bottom-up-only metric can do.
\RSCI reveals methodological gaps in current provider reporting and we set out concrete asks for closing them, so that future bottom-up metrics can be trusted to reflect real-world carbon reductions.
\end{abstract}

\maketitle

\section{Introduction}
\label{sec:intro}

A major challenge in sustainable computing is quantifying the carbon footprint of workloads. 
It is a prerequisite for determining the effectiveness of carbon-reduction strategies, for reporting to regulators and investors, and for making informed decisions about workload scheduling and adaptation. 
Two largely separate approaches exist: bottom-up estimates of operational and embodied carbon, and top-down corporate greenhouse gas (GHG) reporting. % based on provider-wide measurements and downstream allocation to tenants.

\parag{Bottom-up metrics}
Most carbon-aware computing research has focused on optimizing bottom-up metrics like operational carbon: the estimated emissions from a workload's marginal increase in energy consumption. 
This signal is straightforward to obtain (workload energy $\times$ carbon intensity) but does, for example, not capture the embodied carbon of the hardware running the workload.
To provide a more comprehensive metric, the Green Software Foundation proposed the Software Carbon Intensity (\SCI) specification~\cite{sci-spec}, codified as ISO/IEC~21031 in 2024, which expresses the carbon emitted per unit of work and folds in an embodied term.
However, Bashir et al.~\cite{bashir2024} show that \SCI can produce perverse incentives for workload placement and scaling in practice. 
Furthermore, every bottom-up metric, \SCI included, can only capture what a tenant can observe and model.

\parag{Top-down reporting}
Due to regulatory pressure, cloud providers increasingly offer customer-facing dashboards and reports enabling tenants to estimate and disclose their cloud carbon footprint in line with the GHG Protocol~\cite{ghg-protocol-corp}.
While the reporting boundaries and allocation methods can differ significantly by provider, the general methodology is usually public and sometimes even externally audited.
However, the reported numbers are highly aggregated: 
reports mostly contain a single number per month, region, customer, and GHG Protocol scope, sometimes coarsely categorized further into service types like ``Compute'', ``Storage'', and ``Other''.
The reports are published about 14-21 days after the end of each month, which makes them comprehensive but unsuitable for any type of real-time optimization.

\parag{The need for reconcilability}
Today, when an engineering team improves the bottom-up metric of an internal workload, it is entirely unclear whether that improvement reaches the provider's top-down report.
The two can diverge in both directions.
The top-down report contains overheads no bottom-up model can cover without provider-internal data.
And conversely, the levers a bottom-up metric does capture---such as lowering a VM's utilization and thus its power draw---may not move the reported number at all, if the provider allocates carbon to tenants by usage-hours or revenue instead of measured energy.
\todo{maybe switch order and make this more general}

The Green Software Foundation explicitly endorses this separation, distinguishing ``measuring for action'' (per-workload signals that drive engineering) from ``measuring for reporting'' (organization-level inventories that satisfy regulators) and arguing that the two serve different goals and should not be conflated~\cite{gsf-measuring}.
\emph{We disagree.}
Now that regulators require firm-level GHG disclosure~\cite{eu2022csrd,SEC2024_3311275,sbti2024netzero} and carbon pricing is increasingly implemented, a per-workload metric whose gains never land in official reporting will not be funded, sustained, or trusted at scale.


\parag{Contributions}
We argue that bottom-up carbon metrics should \emph{reconcile} with the provider's top-down GHG report by construction.
Towards this, we make the following contributions:

\begin{itemize}\setlength{\itemsep}{1pt}
\item We review the state of the art in cloud carbon accounting, including a survey of 48 providers and a lever-by-provider diagnostic matrix showing which engineering actions actually move which provider's reported number.
\item We introduce \RSCI, which anchors a bottom-up energy estimate to the provider's per-scope, location-based report through a \emph{residual bridge}, so per-workload numbers sum to the reported total by construction.
\item We show that reconciliation constrains only the reported \emph{total} and leaves the per-workload allocation free, so the residual bridge can be allocated to reward sustainable behavior, such as peak-shaving and procurement-deferral, that no bottom-up-only metric can express.
\item We diagnose why \RSCI cannot yet be computed on today's provider reports, and set out concrete asks on provider reporting, customer-side estimation, and standards that would close the gap.
\end{itemize}


%temporal and spatial load shifting~\cite{wiesner2021letswaitawhile,radovanovic2022google,hanafy2024gaia,sukprasert2024limits,gsteiger2024caribou,murillo2024cdn,lechowicz2023pause,lechowicz2024carbonclipper}, elastic scaling~\cite{hanafy2024carbonscaler}, virtual-energy abstractions~\cite{souza2023ecovisor}, and quality- or approximation-side adaptation~\cite{wiesner2025qualitytime,vitali2026adaptive}; even infrastructure planning that takes embodied carbon seriously at fleet scale still optimises the per-workload signal against operational energy~\cite{acun2023carbon_explorer}.

%unprecedented rate~\cite{epri2024powering,bashir2024climate,gaffney2025earth}

% A more comprehensive per-workload metric was introduced in 2021 by the Green Software Foundation: the Software Carbon Intensity (\SCI) specification~\cite{sci-spec}, codified as ISO/IEC~21031 in 2024.
% \SCI extends $E\cdot I$ with an explicit embodied term, the amortised cradle-to-gate carbon of the hardware running the workload --- formalised by Gupta et al.'s architectural carbon model ACT~\cite{gupta2022act} --- so that hardware-procurement choices are reflected alongside operational energy.
% The stated intent is a single per-workload number that ranges from a function call to a datacenter, is real-time sensitive, and is meant to drive engineering optimisation.


\section{State of the Art in Cloud Carbon Reporting}
\label{sec:sota}

Under regulatory pressure~\cite{eu2022csrd,SEC2024_3311275,sbti2024netzero}, cloud providers have moved beyond annual corporate sustainability reports toward customer-facing carbon tools that attribute emissions to individual tenants, enabling Scope-3 accounting for cloud usage.
We surveyed 48 cloud service providers in May~2026 (full enumeration and disclosure classification in Appendix~\ref{app:inventory}): 12 ship a first-party, per-customer dashboard, and most of these publish methodologies detailed enough to evaluate their reporting boundaries and allocation rules.

In this section, we focus on the three largest providers (AWS, GCP, and Azure), which offer comparable dashboards and detailed public methodologies.
Despite differing in detail, all three follow the same two steps: they first measure their total datacenter emissions during a month, then allocate that total down to each tenant.
We walk through both steps, note where the providers diverge, and explain \emph{which tenant actions actually move the reported number}.

\parag{Measuring the datacenter total}
Providers decompose their total datacenter emissions over the reporting period along the GHG Protocol's three scopes.
\emph{Scope~1}, direct on-site emissions such as backup-diesel combustion and refrigerant leakage, is usually negligible for most tenants today, though the increasing deployment of behind-the-meter gas generation at datacenters may change this in future.
\emph{Scope~2} is the datacenter's purchased electricity valued at the region's grid carbon intensity.\footnote{The GHG Protocol mandates dual reporting for Scope 2: \emph{location-based} figures reflect actual grid emissions, while \emph{market-based} figures net out procurement instruments like power purchase agreements (PPAs) and renewable energy certificates (RECs), which in practice often moves the market-based number to near zero. Since engineering teams have no control over the provider's procurement choices, the market-based figure is not actionable for workload optimization, which is why \SCI explicitly excludes it~\cite{sci-spec} and why we consider only the location-based figure.}
\emph{Scope~3} covers amortized cradle-to-gate hardware and, depending on the provider, building embodied carbon, upstream fuel-and-transmission losses (FERA), freight, and end-of-life.
No two of the three providers count the same set of Scope-3 categories (Table~\ref{tab:bigthree-s3} in the appendix), so even the headline totals are not directly comparable.

Customer-facing dashboards typically report emissions at the service level (e.g., for GCP, "Compute Engine," "Networking", ...), which is significantly coarser than the per-workload granularity of a bottom-up metric. A single service can encompass dozens of distinct SKUs, and a single SKU can in turn be consumed by thousands of individual workloads;
none of that internal structure survives into what's actually published. 
What customers see instead is Scope 1, 2, and 3 totals per
$$\text{project} \;\times\; \text{service} \;\times\; \text{region} \;\times\; \text{month},$$
published 14 to 21 days after the period closes.
The monthly cadence and delay is a property of the top-down pipeline: emissions are measured and assured at the facility and corporate level over an accounting period, and also allocation to tenants is usually weighted based on the monthly usage.
Corporate GHG inventories are compiled annually, so monthly reporting is already finer than corporate disclosure requires.

\parag{Allocating to tenants}
Each provider then allocates the datacenter total to tenants in proportion to their share of an \emph{activity unit}, but the units are inconsistent and lean on economics to differing degrees.
AWS runs two regimes at once, physical instance-hours for services with dedicated racks, \emph{equivalent revenue} for the rest, without publishing which service falls on which side~\cite{aws-ccft-model3}.
GCP hides an economic weighting inside a physical-looking unit: it distributes each product's energy across SKUs in proportion to list price before allocating by per-SKU usage such as vCPU-hours~\cite{schneider-mattia-2024}.

None of the big three allocates by measured energy, which is certainly not easy to do within virtualized infrastructures and at scale.
As a result, two identical VMs of the same type and duration (one idle, one fully loaded) are attributed the same emissions.
The allocation also washes out \emph{when} a tenant runs.
None of the three returns a per-tenant temporal signal: AWS and Azure apply a monthly-average carbon intensity, and although GCP computes its location-based number on hourly emission factors, it collapses them into a fleet-wide, energy-weighted per-SKU-and-region average before allocating tenants by monthly usage~\cite{schneider-mattia-2024}.
A tenant that shifts the same work to cleaner hours therefore sees no change in its reported number.

\begin{table}[t]
\centering
\caption{Does a metric \emph{respond} when a tenant pulls a sustainability \emph{lever}? Each of the six levers reduces a workload's \emph{real-world} carbon; the last row is an action that does not. Cells give the metric's response: \faithful~faithful (the number moves with real carbon, upward in the last row), \proxyonly~proxy-only (it moves, but driven by list price or a usage proxy, not carbon), \notrefl~not reflected (no response), \perverse~perverse (the number improves although real carbon does not fall)~\cite{bashir2024}. Bottom-up metrics (\S\ref{sec:background}) and the hyperscalers' top-down tools fail on complementary levers; \RSCI (\S\ref{sec:rsci}) responds faithfully to every row by construction. Provider figures are location-based; Azure exports only the near-zero market-based number, so its region response ($^\ast$) is invisible in practice.}
\label{tab:levers}
\footnotesize
\setlength{\tabcolsep}{4pt}
\renewcommand{\arraystretch}{1.2}
\begin{tabular}{@{}l cc ccc c@{}}
\toprule
& \multicolumn{2}{c}{\textbf{Bottom-up}} & \multicolumn{3}{c}{\textbf{Top-down}} & \textbf{Ours} \\
\cmidrule(lr){2-3}\cmidrule(lr){4-6}\cmidrule(l){7-7}
\textbf{Sustainable lever} & \SCI & \OSCI & AWS & GCP & Azure & \RSCI \\
\midrule
SKU / service selection$^{\dagger}$ & \faithful & \faithful & \proxyonly & \proxyonly & \proxyonly & \faithful \\
Right-sizing / autoscaling & \faithful & \faithful & \faithful  & \faithful  & \faithful  & \faithful \\
Utilization of held resources$^{\dagger}$ & \faithful & \faithful & \notrefl   & \notrefl   & \notrefl   & \faithful \\
Region selection (spatial)               & \faithful & \faithful & \faithful  & \faithful  & \faithful$^\ast$ & \faithful \\
Time-shifting (temporal) & \faithful & \faithful & \notrefl   & \notrefl   & \notrefl   & \faithful \\
Avoid demand peaks                     & \notrefl  & \notrefl  & \notrefl   & \notrefl   & \notrefl   & \faithful \\
\midrule
\multicolumn{7}{@{}l}{\emph{Not sustainable (real carbon rises):}} \\
Shift load to older, amortized hardware & \perverse & \faithful & \notrefl   & \notrefl   & \notrefl   & \faithful \\
\bottomrule
\end{tabular}
\\[2pt]
{\scriptsize\textit{$\dagger$ Individually feasible elsewhere: Scaleway and Oracle's power-based mode allocate by measured energy, so utilization responds (Akamai and T Cloud Public claim the same), and Scaleway derives its SKU factors from LCA data. Pricing tier is normalized out everywhere: a spot and an on-demand VM of the same SKU receive identical emissions.}}
\end{table}

\parag{Which levers move the reported number?}
Given this machinery, the natural operator question is whether pulling a sustainability \emph{lever} (an action that intuitively should reduce a workload's carbon footprint) produces a \emph{response} in a metric's reported number, and whether it responds for the right reason.

Bottom-up estimates and top-down reports are two imperfect observers of the same real quantity, and they fail on complementary levers.
Table~\ref{tab:levers} maps six such levers (plus one deliberately unsustainable action) against both metric families, and the pattern is uneven.
Reducing billed units (right-sizing, autoscaling, switching idle resources off) is the one lever that reliably lowers the number, because allocation is linear in the usage unit.
Choosing a cleaner region also works, since Scope~2 is valued at the regional grid intensity, though the lever is invisible in Azure's default export, which carries only the near-zero market-based figure.
The remaining levers fail in one of two ways.
Raising the utilization of resources a tenant already holds, or shifting load off carbon-intensive hours, does not move the number at all, because neither changes the billed usage units.
Selecting a more efficient SKU or service does move the number, but for the wrong reason: at GCP the change is proportional to list price, not to actual energy~\cite{schneider-mattia-2024}, so a cheaper-but-dirtier SKU can appear greener.
The bottom-up columns (\S\ref{sec:background} defines the metrics) show the mirror image: they respond faithfully to every operational lever but are blind to peak avoidance, and a metric with an embodied term can even respond \emph{perversely}: shifting load onto older, amortized hardware improves \SCI while real carbon rises, because the operational increase is masked by a smaller sunk embodied share~\cite{bashir2024}.

These gaps are not fundamental.
Several smaller providers already reflect what the hyperscalers miss: Scaleway and Oracle's power-based mode allocate compute by measured energy, so an idle and a busy VM diverge, which shows the levers are individually feasible, just not unified in any one tool.
The starkest omission is peak-demand avoidance: shifting load away from system peaks is genuine sustainable behavior, because it defers the next round of hardware procurement, yet no monthly report exposes a peak signal.
Unlike embodied carbon, which is a sunk cost once the hardware is built~\cite{bashir2024}, peak avoidance changes \emph{future} emissions, exactly the kind of incentive a bottom-up-only metric cannot encode.

...


Coverage of Scope 2 accounting methods also differs across providers: AWS and GCP customer dashboards report both location-based and market-based Scope 2, while Azure's Carbon Optimization dashboard reports only market-based Scope 2.

\begin{takeaway}
Not every action that improves a bottom-up metric improves the provider's top-down number, and vice versa.
Right-sizing and region choice move the audited figure faithfully; higher utilization, load-shifting, and peak avoidance do not move it at all; and picking a ``greener'' SKU can move it the wrong way.
An engineering team optimizing a bottom-up metric and an auditor reading the top-down report can therefore reach opposite conclusions about the same workload, the gap \RSCI is designed to close.
\end{takeaway}



\section{The Issue with Bottom-Up Carbon Metrics}
\label{sec:background}

A bottom-up metric estimates a workload's carbon from quantities a tenant can measure without provider-internal data.
The de facto choice in most research papers is operational carbon, which we write as \OSCI (operational software carbon intensity) to match the terminology in Bashir et al.~\cite{bashir2024}:
\[
\OSCI \;=\; \frac{\E \cdot \I}{\R},
\]
the workload's energy $\E$ times its region's carbon intensity $\I$ (in gCO$_2$/kWh), per functional unit of work $\R$ (an API call, a training run, a GB stored per month).
It is straightforward, transparent, and it tracks the two levers a tenant controls directly: energy efficiency and carbon-aware timing or placement.
But it says nothing about hardware efficiency.

To capture the missing hardware lever, the Software Carbon Intensity (\SCI) specification (ISO/IEC~21031~\cite{sci-spec}) adds an embodied term:
\[
\SCI \;=\; \frac{\E \cdot \I \;+\; \M}{\R},
\]
where $\M$ is the workload's amortized share of the hardware's embodied carbon, allocated according to its reserved time and resource share as described by Gupta et al.~\cite{gupta2022act}.\todo{maybe a sentence that SCI is intentionally vague about PUE etc., which can be included, but that it cannot observe everything. maybe add this to the last parag}

However, Bashir et al.~\cite{bashir2024} show that \SCI is subject to a \emph{sunk-carbon fallacy}: because \M is amortized flat over the hardware's expected lifespan, \M/\R does not respond to declining hardware efficiency. As a server ages and its energy use \E rises, the embodied component (a sunk cost) remains unchanged, so a scheduler minimizing \SCI can drift toward older, less efficient hardware, increasing total carbon emissions.
They therefore recommend sticking to \OSCI for operational decisions such as scheduling and scaling. 
We argue, however, that \OSCI is incomplete: it does not reward reserving less hardware (beyond the energy savings), purchasing lower-embodied hardware, or avoiding peak-hour demand, which ultimately determines when additional capacity must be procured.

Moreover, independent of the embodied term, \emph{any bottom-up-only metric is structurally blind to provider-side overheads} that are not explicitly modeled, and therefore cannot reconcile to the provider's top-down GHG report.
This means that improving \SCI or \OSCI does not necessarily move the audited number, and vice versa.
\todo{it could be made clearer why this is an issue: bottom-up-only are structurally insufficient to reconcile to top-down, and that this is a problem for trust and adoption. We can also give provider side overhead examples here: PUE, idle and over-provisioned capacity, upstream fuel-and-transmission losses, building embodied carbon, refrigerant leakage, water, and end-of-life}

\todo{Maybe flip Secs 2 and 3: first show the SOTA in cloud carbon reporting, then show the gaps that bottom-up metrics cannot close.}




\iffalse % ===== OLD SEC 3 --- preserved for salvage (provider quotes -> Sec 6, Scope-3 coverage table -> appendix) =====

\paragraph{Customer-facing tools are rare.}
We surveyed 48 cloud providers in May~2026, covering hyperscalers, mid-tier and regional IaaS, CDN/edge, AI-specialised neoclouds, application-PaaS, and sustainability-positioned regional operators (full enumeration with disclosure classification in Appendix~\ref{app:inventory}).
Only twelve ship a first-party, per-customer carbon dashboard --- the bar for a customer to source a GHG-Protocol-compliant Scope-3 Cat-1 input from their cloud spend.\footnote{We exclude marketplace third-party SaaS such as Huella Cloud on Huawei Cloud (vendor: SGS Solutions), since the per-provider allocation is not auditable against the provider's own emission factors. Exoscale's calculator is borderline (Exoscale-branded surface, CloudAssess engine by Resilio/Kleis) and is omitted.}
Of the remaining 36, fourteen publish corporate-level Scope~1/2/3 disclosures without any per-customer surface; the other 22 publish only marketing-grade renewable/PUE claims or nothing at all.
Most starkly, the fastest-growing segment --- AI/GPU clouds (CoreWeave, Lambda, Crusoe, Nebius, RunPod, Together, Replicate, Hyperstack, Verda, Fireworks~AI, Paperspace) --- has \emph{zero} per-customer tools.
Compute capacity for AI training and inference is being built at the fastest pace in cloud history with the \emph{least} customer carbon transparency.
The rest of this section focuses on the twelve per-customer tools.

\paragraph{One shared template, four axes of variation.}
Strip the marketing and every customer tool computes the same shape:
\[
\mathrm{customer\,carbon} \;=\;
\textstyle\sum_{r}\, u_{r}\,\cdot\,f_{r} \;+\; \mathrm{overheads},
\]
where $u_{r}$ is an activity unit (vCPU-hours, instance-hours, bytes, CPU-time, GB-month, \$-spend, server-hours) and $f_{r}$ is a per-unit factor.
Provider variation lives on four orthogonal axes: (A1)~which activity unit, (A2)~how $f_{r}$ is derived, (A3)~embodied carbon treatment, (A4)~idle/overhead allocation.
Eleven of the twelve customer tools differ on the two most discriminating axes (A1, A2); T Cloud Public is the twelfth (launched 2026-02-23) but has not disclosed its per-resource allocation rule.
The prior literature's three-family taxonomy (usage-time / traffic / component-LCA) puts AWS-foundational and OVHcloud-dedicated in different families even though both reduce to $\sum u_{r} f_{r}$, and we replace it with the per-axis decomposition below.

% [DROPPED 2026-06-26] Former Figure 1 (axes A1xA2 mapping) removed --- unclear payload.
% Preserved here in case the axis decomposition is reused in the rewritten Sec 3.
% \begin{figure}[t]
% \centering
% \footnotesize
% \setlength{\tabcolsep}{4pt}
% \renewcommand{\arraystretch}{1.15}
% \begin{tabularx}{\columnwidth}{@{}l L L L L@{}}
% \toprule
% & \multicolumn{4}{c}{\textbf{A2 --- per-unit-factor source}} \\
% \cmidrule(l){2-5}
% \textbf{A1 --- unit} & price/rev. & LCA $\times$ inv. & meter $\rightarrow$ bucket & benchmark \\
% \midrule
% vCPU-h        & GCP-cust    &              & AWS-FS, Azure, IBM, Ali & \\
% instance-h    & AWS-nFS     &              &                         & \\
% CPU-time      &             &              &                         & Fastly-C \\
% bytes         &             &              & Akamai$^{\dagger}$       & CF, Fastly-D \\
% GB-month      &             &              & (same as compute)       & \\
% server-h      &             & OVH, Scaleway$^{\ddagger}$ &              & \\
% \$-spend      & AWS-nFS, Oracle-sp, Azure'21 & &                      & \\
% \bottomrule
% \end{tabularx}
% \\[2pt]
% {\scriptsize\textit{$\dagger$ Akamai claims per-server kWh metering $\times$ byte/machine-utilization share.}}\\
% {\scriptsize\textit{$\ddagger$ Scaleway's monthly report combines per-component LCA with actual measured CPU consumption.}}
% \caption{Per-customer carbon tools on the activity-unit (A1) $\times$ factor-source (A2) axes.}
% \label{fig:axes}
% \end{figure}

\paragraph{The ideal weighting is measured energy --- almost no one delivers it.}
The right question is not which usage unit a provider uses but how well that unit tracks the customer's actual energy draw.
For network, bytes are approximately linear in energy, so byte-share allocation is structurally sound (Cloudflare, Fastly-Delivery, Akamai's byte component).
For compute the relationship is highly non-linear --- an idle server still draws on the order of half its full-load power --- so vCPU-hours, instance-hours, and \$-spend systematically misweight idle and busy workloads.
Four customer-surface allocations approach energy-direct, in decreasing concreteness: \emph{Scaleway} combines per-component LCA with actual measured CPU consumption and published Boavizta consumption profiles for the monthly report, so an idle and a fully-loaded VM at the same SKU receive different allocations;\footnote{\url{https://www.scaleway.com/en/docs/environmental-footprint/}; per-product calculation pages compiled in our reference set.}
\emph{Akamai} meters kWh per server and per network device in its colocations\footnote{Akamai 2025 emissions-reporting policy and customer-reporting blog series.} and allocates DC-total energy by ``byte utilization and machine utilization'';
\emph{Oracle}'s power-based mode tracks kWh by service workload (LBM~+~MBM) but does not disclose the shared-hardware allocation algorithm;\footnote{\url{https://docs.oracle.com/en-us/iaas/releasenotes/governance/carbon-analysis-powercalculation.htm}, June~2025.}
and \emph{GCP-internal} measures power per machine via Borg telemetry but deliberately re-projects to vCPU-h~$\times$~list-price at the customer surface~\cite{schneider-mattia-2024}.
The remaining compute-allocation tools (AWS, Azure, IBM, Alibaba, OVHcloud-currently) propagate a usage-unit proxy: an idle and a fully-loaded VM of the same SKU and duration produce the same number.
T Cloud Public claims a utilization-share allocation from measured device-level power but has not yet published its per-resource formula, so the customer-side weighting cannot be independently checked.

\paragraph{Coverage and verification are shallower than the dashboards suggest.}
Table~\ref{tab:bigthree-s3} breaks the Big-3 Scope-3 perimeter down by GHG-Protocol category.
No two providers cover the same set: AWS includes FERA and full Cat~2 capital goods (with buildings); GCP adds business travel and employee commuting but excludes end-of-life; Azure is the only Big-3 to include Cat~12 end-of-life yet excludes both FERA and building embodied carbon.
None publishes the materiality assessment the GHG Protocol requires to justify a category exclusion.
Headline numbers are therefore not directly comparable.
Third-party assurance is similarly shallow: AWS's Apex statement covers Scope-3 Cat~2/3/4 \emph{methodology} only, GCP's Fraunhofer review covers the LCA methodology only, Azure's WSP statement covers the Scope-3 \emph{document} only; \textbf{no Big-3 customer tool has been audited end-to-end}, and only Alibaba's Energy~Expert carries third-party \emph{data} assurance (Bureau Veritas, ISO~14064-3) of the per-customer numbers.
The temporal surface is similarly thin: \textbf{only Scaleway and Fastly publish daily customer-facing data}; the other nine, including all of the Big-3, are monthly; \emph{no} customer tool publishes hourly customer-allocated emissions, so any carbon-aware-scheduling claim today rests on signals outside the customer tool itself.

\paragraph{The internal/external gap is structural --- and is the \RSCI residual.}
Three providers say this out loud.
Microsoft's 2021 Scope-3 transparency whitepaper, attached to its own description of the customer-facing tool, states that ``\emph{the underlying methodologies and emissions findings generated from the calculator will differ from those reflected in Microsoft's corporate disclosure}''~\cite[p.~2, fn.~2]{microsoft-s3-2021}, and the January~2026 Azure methodology page repeats the point in operational terms: ``\emph{usage for emissions calculations might not match your Microsoft usage for billing purposes}''~\cite{azure-methodology-2026}.
GCP measures machine-level power via Borg hourly but re-projects to a monthly \emph{list-price-proportional} SKU factor at the customer surface, and Schneider \& Mattia note (\S3.6) that ``\emph{customers do not receive a signal about how switching from SKU~A to SKU~B can reduce their actual emissions, but rather their allocated emissions based on list prices}''~\cite{schneider-mattia-2024}.
AWS's Model~3.0 defines a foundational/non-foundational service split but publishes neither the service list nor the share of emissions allocated by equivalent revenue~\cite{aws-ccft-model3}.
The pattern is not three isolated quirks: every customer tool ships a methodologically distinct, thinner subset of what the provider measures internally, and the size of that gap is the residual that \S\ref{sec:rsci} formalises through \RSCI.
Bottom-up engines such as BoaviztAPI~\cite{simon2024boaviztapi} and Cinergy~\cite{jacquet2025cinergy} reach the same verdict from the opposite side and are the natural feed for the bottom-up term in \RSCI; they do not themselves reconcile to the provider report.

\fi % ===== END OLD SEC 3 =====

\section{Reconciling Bottom-Up and Top-Down}
\label{sec:rsci}

% NOTE: framing wrapper depends on §5 experiments. Math is committed; opening list
% of "three gaps closed" follows the §2 critique structure.

We introduce \emph{Reconcilable \SCI} (\RSCI), a class of actionable, per-workload metrics designed to close three gaps:
\begin{enumerate}
  \item the sunk-carbon fallacy of \SCI
  \item the loss of embodied-carbon incentives in \OSCI
  \item the disconnect between every bottom-up-only \SCI variant and the cloud provider's top-down carbon report
\end{enumerate}
\RSCI does so by anchoring a bottom-up energy model to the provider's per-scope top-down report via residual bridges, and by allocating each residual component across workloads with a weight chosen to match its physical driver.

\parag{Setup}
Let $i$ index \emph{workload instances}: one evaluation of a workload over period $t_i$. 
The provider issues a top-down report broken into Scope-1, Scope-2, and Scope-3 totals at the level of a \emph{reporting slice} indexed by service $s$ and region $r$ (the finest granularity providers publish: per-month, per-project, per-service).
\todo{Revise notation.}
Let $C_{s,r}^{\downarrow\text{S1}}$, $C_{s,r}^{\downarrow\text{S2}}$, and
$C_{s,r}^{\downarrow\text{S3}}$ denote the reported per-scope emissions of slice $(s,r)$.

\parag{Bottom-up energy}
For workload $i$ in slice $(s,r)$ we estimate energy from per-resource quantities:
\begin{equation}
E_{i,s,r} \;=\; \sum_{p} q_p \cdot \varepsilon_p,
\label{eq:bu-energy}
\end{equation}
where $p$ indexes resource types (vCPU-hours, GPU-hours, GB-hours of storage, I/O ops, network GB) and $\varepsilon_p$ is a per-type energy coefficient.
Concrete instantiations of $\varepsilon_p$ are available from open-data projects such as BoaviztAPI~\cite{simon2024boaviztapi} (per-component LCA factors and CPU power curves) and per-VM deterministic models such as Cinergy~\cite{jacquet2025cinergy} (production-validated on OVHcloud). \RSCI is orthogonal to the choice of bottom-up engine.
Unobservable resource types simply enlarge the residuals defined below.

\parag{Residual bridges}
For each top-down component $k$ reported by the provider, the residual is the gap between what the provider reports and what the bottom-up model accounts for:
\begin{equation}
\Delta^{k}_{s,r} \;=\; C^{\downarrow k}_{s,r} \;-\; O^{k}_{s,r},
\label{eq:residual-general}
\end{equation}
where $O^{k}_{s,r}$ is the bottom-up estimate of that same quantity (zero for components the bottom-up model does not address).
At the coarsest decomposition (one residual per scope)
\begin{equation}
\Delta^{\text{S1}}_{s,r} = C^{\downarrow\text{S1}}_{s,r}, \;\;
\Delta^{\text{S2}}_{s,r} = C^{\downarrow\text{S2}}_{s,r} - \!\sum_i\! E_{i,s,r} I_r, \;\;
\Delta^{\text{S3}}_{s,r} = C^{\downarrow\text{S3}}_{s,r},
\label{eq:residual-scopes}
\end{equation}
with $I_r$ the location-based grid intensity of region $r$.
$\Delta^{\text{S1}}$ absorbs diesel and refrigerant emissions; $\Delta^{\text{S2}}$ absorbs PUE, idle capacity, market-vs-location adjustments, and energy-model error; $\Delta^{\text{S3}}$ absorbs embodied hardware, building embodied, FERA, upstream transport, and the other Scope-3 categories of \S\ref{sec:sota}.
Where the provider reports finer-grained sub-categories (and where the operator finds it useful) the same definition (Eq.~\ref{eq:residual-general}) admits further splitting; we will use a minimal split of $\Delta^{\text{S3}}$ into FERA, embodied IT, and non-IT (building, freight, end-of-life) below.

\parag{Reconciliation}
Reconciliation is performed per slice $(s,r)$.
Each residual component $k$ is allocated across workloads by a weight $w^{k}_{i,s,r}$ satisfying $\sum_i w^{k}_{i,s,r} = 1$, and the per-workload metric is
\begin{equation}
\RSCI_i \;=\; \frac{1}{R_i}\sum_{s,r} \Bigl[\, E_{i,s,r}\cdot I_r \;+\; \sum_{k} w^{k}_{i,s,r}\cdot\Delta^{k}_{s,r} \,\Bigr].
\label{eq:rsci}
\end{equation}
The first inner term is the bottom-up operational footprint; the second is the workload's share of each residual component under the chosen weight.
The choice of $w^{k}$ is left open by Eq.~\ref{eq:rsci}: different weight families yield metrics that incentivise different operator behavior, while preserving reconciliation by construction (Eq.~\ref{eq:reconcile}).
The next subsection makes this design surface explicit.
\todo{shorten}

\subsection{Allocating the Residual}
\label{sec:rsci-design}

Three weight families cover the residual components today's provider reports
expose, each motivated by a different physical driver of the residual.

\paragraph{Energy-share weights.}
\begin{equation}
w^{k}_{i,s,r} \;=\; \frac{E_{i,s,r}}{E_{s,r}}, \qquad E_{s,r} = \sum_i E_{i,s,r}.
\label{eq:w-energy}
\end{equation}
Appropriate when the physical driver of $\Delta^{k}$ is electricity drawn: the load-proportional portion of $\Delta^{\text{S2}}$ (PUE overhead, MBM adjustments, energy-model error), and the FERA sub-component of $\Delta^{\text{S3}}$ (upstream fuel and T\&D losses are physically proportional to kWh consumed).

\paragraph{Static-share weights.}
\begin{equation}
w^{k}_{i,s,r} \;=\; \frac{1}{M_{s,r}},
\label{eq:w-static}
\end{equation}
where $M_{s,r}$ is the number of active workloads in the slice.
Appropriate when the driver is facility uptime or facility existence --- costs incurred by being ``in the room'' rather than by consumption.
Sharma and Fuerst's accountable-FaaS framework~\cite{sharma2024accountable_footprint_serverless} adopts this rule for idle power and for embodied carbon under a single-server abstraction, on the principle (their \S5) that embodied is more naturally treated as a \emph{static sunk cost} than a usage-based cost.
We adopt the same rule for $\Delta^{\text{S1}}$ (diesel backup, refrigerant leakage) and for the non-IT sub-components of $\Delta^{\text{S3}}$ (building embodied, freight, end-of-life).

\paragraph{Peak-attribution weights.}
\begin{equation}
w^{k}_{i,s,r} \;=\; \frac{\varphi_i}{\sum_j \varphi_j},
\label{eq:w-peak}
\end{equation}
where $\varphi_i$ is workload $i$'s contribution to fleet peak --- formally, the Shapley value~\cite{shapley1953value} of $i$ in the cooperative game whose value function is the slice's peak provisioned capacity.
Appropriate for the embodied-IT sub-component of $\Delta^{\text{S3}}$, whose physical driver is peak provisioning: the workload at peak forces the next server purchase; the workload at the trough does not.
The same physical argument was recently made for embodied-carbon attribution within a single server by Han et al.~\cite{HotCarbon2024GameTheoretic}, who frame ``\emph{embodied carbon attribution as the problem of demand-driven supply provisioning for hardware}''~\cite[\S3.4]{HotCarbon2024GameTheoretic}; \RSCI generalises the move to the slice / fleet scope via the top-down residual.
Exact Shapley over the peak-capacity game is exponential in workload count and infeasible online; following the precedent set by~\cite{sharma2024accountable_footprint_serverless,HotCarbon2024GameTheoretic}, we treat $\varphi_i$ as a \emph{best-effort Shapley approximation} satisfying the four Shapley properties (efficiency, null-player, symmetry, linearity).
Practical proxies that retain the qualitative ordering include: (a)~\emph{peak-hour energy share} $\varphi_i = E_i(t^{*})$ with $t^{*}$ the slice peak hour; (b)~\emph{leave-one-out marginals} $\varphi_i = \text{peak}(\mathcal{N}) - \text{peak}(\mathcal{N}\setminus\{i\})$; (c)~\emph{sampled-permutation} Shapley over a few hundred coalition orderings~\cite{castro2009polynomial}.
The choice of proxy is itself a design surface; \RSCI reconciles under any choice satisfying $\sum_i w^{k}_{i,s,r} = 1$.

\paragraph{Default mapping.}
Mapping the three families to the residual components of Eq.~\ref{eq:residual-scopes} (and its finer Scope-3 split) yields a default schema: energy-share for $\Delta^{\text{S2}}$ and $\Delta^{\text{S3-FERA}}$; static-share for $\Delta^{\text{S1}}$ and the non-IT sub-components of $\Delta^{\text{S3}}$; peak-attribution for $\Delta^{\text{S3-embodied-IT}}$.
This is a starting point, not a prescription: operators with finer telemetry can split residual components further or substitute other weights, and \RSCI reconciles as long as each chosen weight family sums to one across workloads in the slice.
The vision contribution of this paper is the \emph{recognition that the residual is a design surface for sustainable-behavior incentives}, not a fixed allocation rule.

\paragraph{Reconciliation by construction.}
Because $\sum_i w^{k}_{i,s,r} = 1$ for every component $k$, summing $\RSCI_i \cdot R_i$ across all workloads in a slice yields the top-down per-scope total:
\begin{equation}
\sum_i \RSCI_i \cdot R_i \;=\; \sum_{s,r}\Bigl(C^{\downarrow\text{S1}}_{s,r} + C^{\downarrow\text{S2}}_{s,r} + C^{\downarrow\text{S3}}_{s,r}\Bigr).
\label{eq:reconcile}
\end{equation}
Reconciliation is not an empirical property to verify; it is a structural identity that holds under \emph{any} admissible weight choice for \emph{any} residual component.

\paragraph{Reconciliation pins the sum, not the allocation.}
Eq.~\ref{eq:reconcile} pins the \emph{fleet total} to the audited number, but it does not pin the \emph{per-workload allocation}, which the weights $w^k$ remain free to shape.
This is what lets \RSCI reward behavior the top-down report is blind to.
Consider temporal load-shifting, and read the operational term at the resolution the tenant actually observes: $E_{i,s,r}\cdot I_r$ abbreviates $\sum_t E_i(t)\,I_{r,t}$, with $I_{r,t}$ the \emph{hourly} location-based grid intensity.\footnote{We keep the coarser $I_r$ in the equations above for readability; propagating the hourly index through them is left to a notation pass.}
Suppose workload $i$ moves fixed energy $E_i$ from an hour of intensity $I_0$ to a cleaner hour $I_1 < I_0$.
Its operational term falls by $E_i(I_0-I_1)$; the provider's $C^{\downarrow\text{S2}}_{s,r}$ is a monthly average and does not move, so $\Delta^{\text{S2}}_{s,r}$ rises by the same amount.
Under energy-share weights (Eq.~\ref{eq:w-energy}) the shifter reabsorbs only its own share $E_i/E_{s,r}$ of that increase, for a net change in attributed carbon of $-\,E_i(I_0-I_1)\bigl(1-E_i/E_{s,r}\bigr) < 0$.
Load-shifting therefore lowers $\RSCI_i$ for every workload smaller than the whole slice: the incentive survives anchoring.

What does \emph{not} change is the slice sum, which by Eq.~\ref{eq:reconcile} equals the time-blind provider report and stays fixed no matter how much load the operator shifts.
This is deliberate, not a defect.
\RSCI does not manufacture a fleet-level reduction the audited books do not show; doing so would break reconciliation and, with it, the trust that motivates the metric (\S\ref{sec:intro}).
The absolute reduction is not lost but relocated: it appears as the growth of $\Delta^{\text{S2}}_{s,r}$, a measured quantity of how much clean-hour behavior the provider's coarse methodology declines to credit.
\S\ref{sec:practice} reads this residual as a diagnostic of provider reporting.

\paragraph{Properties.}
Under the default mapping, \RSCI closes the three gaps from the section opening as direct consequences of the chosen weights.
(i) \emph{Sunk-carbon}: \RSCI has no per-server $M$ term.
Embodied carbon lives in $\Delta^{\text{S3-embodied-IT}}$ and is allocated by the peak-attribution weight (Eq.~\ref{eq:w-peak}).
A workload on older, less efficient hardware draws more energy through $E_{i,s,r}\cdot I_r$ in Eq.~\ref{eq:rsci} and reports a higher \RSCI regardless of how embodied is amortised on the provider's books --- the operational term alone carries the sunk-carbon-avoidance argument.
(ii) \emph{Embodied incentive, and with it peak-shaving and procurement deferral}: under static-share or peak-attribution weights for the embodied component, workloads sized to fit actual demand pay less embodied than over-reserved ones --- a response \OSCI structurally cannot supply.
Under peak-attribution weights (Eq.~\ref{eq:w-peak}) specifically, a workload running at the slice peak draws a larger embodied share than one running at the trough, \emph{even at identical energy} --- the incentive structure operators need to defer fleet expansion.
No bottom-up metric (\SCI, \OSCI, \TSCI) can supply it, because the design surface --- a residual to allocate --- does not exist without top-down anchoring.
(iii) \emph{Reconciliation}: Eq.~\ref{eq:reconcile}.

\paragraph{Negative residuals are admissible.}
Eq.~\ref{eq:residual-general} does not constrain the sign of $\Delta^{k}_{s,r}$.
When the bottom-up estimate $O^{k}_{s,r}$ exceeds the provider's reported $C^{\downarrow k}_{s,r}$ for a component, the residual is negative and the corresponding term in Eq.~\ref{eq:rsci} \emph{reduces} the workload's attributed footprint.
Empirically this case arises: Jacquet et al.~\cite{jacquet2025cinergy} show that current provider attribution can under-estimate operational CO$_2$ by up to 3$\times$ for high-utilisation VMs.
\RSCI absorbs such cases by construction --- reconciliation (Eq.~\ref{eq:reconcile}) still holds --- and a negative $\Delta^{k}_{s,r}$ is itself a diagnostic: it signals that the provider's top-down report under-counts what a credible bottom-up model attributes to the slice.
The metric captures the top-down report; whether the top-down report captures physical truth is the provider's accountability, not the metric's.

% TODO (2026-05-19): refresh the Real-time estimation block below under the
% per-component weight schema. The xi = mu*I_r + rho collapse assumed energy-share
% allocation for ALL residuals; under the new schema, different components ride on
% different drivers (electricity, fleet existence, peak share), so a single per-Wh
% effective intensity no longer captures the metric. A real-time estimator under the
% new schema needs per-component projection rules (mu only for the energy-share
% components; static-share and peak-attribution components need their own state
% estimation). Block commented out below; restore + rewrite when ready.
%
% \paragraph{Real-time estimation.}
% The retrospective definition requires $\Delta$ and $E_{s,r}$ from the provider's
% delayed report. To use \RSCI as a real-time optimisation signal we factor out grid
% intensity:
% \[
% \xi_{s,r} \;=\; \mu_{s,r}\cdot I_r \;+\; \rho_{s,r},
% \]
% where $\mu_{s,r} = C_{s,r}^{\downarrow\text{S2}}/(E_{s,r}\cdot I_r)$ is the
% \emph{Scope-2 multiplier} (dimensionless; captures PUE and energy-model error) and
% $\rho_{s,r} = (C_{s,r}^{\downarrow\text{S1}}+C_{s,r}^{\downarrow\text{S3}})/E_{s,r}$
% is the \emph{non-electricity overhead intensity} (Scope~1 + 3 per Wh). Both terms are
% load-independent and slow-moving; given estimates $\hat{\mu}$, $\hat{\rho}$ from the
% last reconciled month,
% \[
% \widehat{\RSCI}_i = \frac{1}{R_i}\sum_{s,r} E_{i,s,r}\bigl(\hat{\mu}_{s,r}\cdot I_r + \hat{\rho}_{s,r}\bigr),
% \]
% with current $I_r$. With no history, $\hat{\mu}=1$, $\hat{\rho}=0$, and
% $\widehat{\RSCI}$ degenerates to \OSCI --- the framework degrades
% gracefully to the prior-art operational metric when reconciliation data is
% unavailable.

\paragraph{Precision scales with data fidelity.}
\RSCI is a \emph{class} of metrics parameterised by the choice of weight family per residual component, not a fixed equation.
Its precision is bounded by three axes: bottom-up model quality (coverage and accuracy of $\varepsilon_p$), top-down reporting granularity (slice resolution + cadence), and the unit-of-work $R$.
As each improves, $\Delta$ shrinks and the residual-allocation choice matters less; reconciliation (Eq.~\ref{eq:reconcile}) holds at any level of data fidelity and under any admissible weight choice.
The next section confronts the theory with today's provider reports.

\section{Evaluation}
\label{sec:demo}

\outdatedbanner
\begingroup\color{blue}

% Outline only (2026-06-06); prose TBW. Story + structure agreed --- one figure per gap.
% Controlled example: real load + real grid, KNOWN residual Delta (the ground truth no
% provider exposes). Report direction/ordering, not magnitudes (Delta is a knob).

\S\ref{sec:rsci} argued that \RSCI closes the three gaps of \S\ref{sec:background} and \S\ref{sec:sota} in theory.
We demonstrate each on one controlled example: a two-SKU fleet (an amortised A100 node and a new H100 node) serving a real LLM-inference trace on a real grid-intensity trace, with a \emph{known} slice residual $\Delta^{\text{S3}}$.
The residual magnitude is an illustrative input --- the quantity no provider exposes, and the reason the example must be controlled --- so we report direction and ordering, not absolute magnitudes; only the two measured \emph{shapes} (the load swing and the grid-intensity swing) drive the qualitative results.

\paragraph{Gap 1 --- no sunk-carbon fallacy.}
\todo{Fig: per-token embodied charge by SKU. The fully-amortised A100
($\text{age}=\text{lifespan}$) carries $M=0$ under \SCI, so \SCI rates the worn GPU the
\emph{cleanest}; \RSCI routes embodied through the residual (worse efficiency $\to$ more
energy $\to$ more residual share), so it is structurally fallacy-free.}

\paragraph{Gap 2 --- the embodied/peak incentive.}
\todo{Fig: per-token carbon vs.\ fleet utilisation. \OSCI and \SCI are flat
($\sim$1.1$\times$ idle$\to$busy); \RSCI under peak-share rises ($\sim$7$\times$). Only
\RSCI makes a busy fleet cost more per token; the energy-share vs.\ peak-share slope is the
design surface (\S\ref{sec:rsci-design}). No bottom-up metric produces this response.}

\paragraph{Gap 3 --- reconciliation to the top-down total.}
\todo{Fig: stacked attributed-carbon breakdown. \OSCI captures operational only;
\SCI adds active-server embodied; \RSCI closes to the constructed provider total by
construction (Eq.~\ref{eq:reconcile}). The bottom-up metrics structurally see only a
minority of the reported number.}

\endgroup

\section{Why \RSCI Cannot Be Computed Today}
\label{sec:practice}

\outdatedbanner
\begingroup\color{blue}

% ===== SEC 6 SCAFFOLD (2026-07-03) — argument skeleton only; prose TBW in a later pass.
% Terminology (lever / response / residual / weight) and Table 1 in Sec 2 are NOT final,
% so keep references to them abstract until they lock. Case-study numbers are from prior
% internal analyses and must be re-derived in one consistent pass before camera-ready.
% Argument arc: conditional promise -> what recovery requires -> evidence -> diagnostic -> bridge.

\paragraph{Setup: the \RSCI column is a conditional promise.}
\S\ref{sec:demo} showed \RSCI closes the gaps when the residual is \emph{known}, and Table~\ref{tab:levers} claims \RSCI responds faithfully to every lever by construction.
That claim is conditional: it holds only if the operator can recover the residual from the report \emph{and} distribute it across workloads by an appropriate weight, both of which require the report to expose each residual component's driver.
This section asks whether today's reports meet that condition; drawing on \S\ref{sec:sota}, the prior is no.

\begin{itemize}
\item \textbf{What recovery requires.} State the criterion as one requirement per weight family (\S\ref{sec:rsci-design}): the residual must be \emph{decomposed} into components; it needs a \emph{physical} (energy-based) allocation basis, or the energy-share weight cannot be formed; and it needs \emph{resource-level, billing-frequency} granularity, or the peak and temporal weights have nothing to attach to. Today's report supplies none of the three cleanly.
\item \textbf{Evidence: real exports, each defeating one requirement.}
  \begin{itemize}
    \item \emph{AWS Frankfurt} --- allocation basis: most of the reported total sits in the revenue-allocated ``Other'' bucket, an undifferentiated lump with no physical driver to weight against. \todo{re-derive the $\sim$70\,\% figure in the consistent pass.}
    \item \emph{Azure Frankfurt} --- the anchor itself: exported location-based Scope-2 is near zero, so the quantity the operational term reconciles against is missing and the region and time levers leave no recoverable response. \todo{re-derive.}
    \item \emph{Scaleway} --- partial existence proof: daily granularity \emph{exists} (the levers are individually feasible), but the daily number is flat under load and idle\,$\approx$\,stress is confounded, so the \emph{union} \RSCI needs is unrealised. \todo{data pending; keep qualitative until the re-run lands.}
  \end{itemize}
\item \textbf{Diagnostic read-out.} Line the failures up against the not-reflected cells of Table~\ref{tab:levers} (utilization, time-shifting, peak): \RSCI could carry a faithful response to each, but the report exposes nothing of the matching shape, so the residual --- not the workload --- absorbs the lever. The residual is not merely large but \emph{undifferentiated}, and its size measures how much of each lever's real-carbon effect the provider declines to expose. This is \RSCI as a diagnostic.
\item \textbf{Close (bridge to \S\ref{sec:asks}).} The barrier is structural (coarse, delayed, opaque) but not fundamental: Scaleway and Oracle's power-based mode show the pieces are individually feasible, so it is a reporting \emph{choice}, which \S\ref{sec:asks} takes up.
\end{itemize}

\endgroup

\section{Call to Action}
\label{sec:asks}

Reconciliation by construction is structurally available, yet \S\ref{sec:practice} shows it is empirically unrealised on today's reports.
Closing that gap is a shared task for providers and standards bodies, and it opens further roles for the metric.

\paragraph{What providers must report.}
Customer carbon must be reported at \emph{billing frequency} and resource granularity, so that the temporal and utilization levers of Table~\ref{tab:levers} land in the number rather than the residual.
Providers already hold the data: GCP computes its location-based figure on hourly emission factors before collapsing them to a monthly average (\S\ref{sec:sota}), so finer reporting is a disclosure choice, not a measurement problem.
Revenue and ``equivalent-revenue'' fallbacks (AWS non-foundational; \S\ref{sec:sota}) should give way to physical-telemetry allocation, and each customer tool should reconcile with the provider's corporate disclosure: Microsoft's open admission that the two do not match (\S\ref{sec:sota}) is a defect against the customer's reporting need, not a feature.
Excluded Scope-3 categories require the published materiality assessments the GHG Protocol already mandates.
We do not ask providers to expose proprietary telemetry, only to report often enough that customer-side estimation, such as cross-tenant pooling of methodology-driven overheads and state-estimation over reconciled-month sequences, can converge on the residual.

\paragraph{What standards must fix.}
The GHG Protocol is under-specified for cloud~\cite[see][]{schneider-mattia-2024}: shared-infrastructure allocation, equipment lifetimes, temporal matching of energy attributes, and building embodied carbon are interpretive gaps in the standard, not violations~\cite[see ``Under-Specified GHG Protocol Areas'']{bashir2024}.
A cloud-specific gold standard should close these and require independent assurance of the \emph{tool implementation}, not just the methodology document.

\paragraph{Outlook.}
Two roles beyond operator scheduling follow directly from reconciliation (Eq.~\ref{eq:reconcile}).
\RSCI is the only metric in its family that can be \emph{honestly reported to end users}, since it reflects the embodied, peak, and overhead costs a request actually triggers rather than operational electricity alone; a per-(request-class, submission-time) view marginalised over spatial placement keeps this routing-invariant while preserving the user's temporal and sizing levers.
It is likewise usable for a customer's own \emph{Scope-3 Cat-1} accounting of cloud use, attributing the provider's disclosed total across internal teams at workload granularity with the guarantee that the parts sum to the whole.
The decomposition is also \emph{signal-agnostic}: swapping the carbon report for a provider's water disclosure yields a reconcilable ``water per unit of work'' metric, and Bashir's \TSCI could serve as the provider-side anchor for \RSCI on the customer side.

\todo{open problems}
allocating the residual across tenants sharing serverless warm pools.
bounding reconciliation error once the peak-attribution weight $\varphi_i$ is estimated rather than computed exactly, which would upgrade the guarantee from ``reconcilable by construction'' to provably $\varepsilon$-close under practical attribution.

% Do not yet touch these
\todo{Tenant vs customer terminology}
\todo{explain MBM and LBM, and why they are relevant to the residual.}
\todo{GHG vs carbon vs CO2e terminology.}
\todo{current carbon accounting is imperfect and does not always incentivize the right downstream behavior, and the standards are under continuous revision. That is a separate problem. \RSCI is defined against the \emph{structure} of GHG-Protocol-style accounting, not any particular vintage, so future revisions slot into the same framework. add citations for the critique of GHG-Protocol-style carbon accounting (market-based instruments, RECs/PPAs, location- vs. market-based divergence, ongoing Scope 2 revision, etc.)}
% 2. Cuts I made in the §7 merge (so you can veto any): the (a)(b)(c) Kalman/pooling enumeration collapsed to one clause; the \hat\mu/\hat\rho references dropped entirely (they were dangling anyway, only defined in the commented-out real-time block); the "Scaleway/OVHcloud already publish water" detail and the energy-coefficient-standardization future-work item removed; the routing-invariance mechanism (the φ_i marginalization argument) compressed to just the claim. The bounded-correlation open problem survived as a paper-strengthener.
% 3. Two references now lean on things not yet solid: the §4 footnote points at a "notation pass" (your existing \todo{Revise notation} at the §4 setup), and §7's provider ask reuses the GCP-collapse claim from §2 line 181 — both already sourced, just noting the coupling.


\bibliographystyle{ACM-Reference-Format}
\bibliography{references}

\appendix
\onecolumn

\section{Cloud Provider Survey}
\label{app:inventory}

All 48 cloud providers reviewed for §\ref{sec:sota}, clustered by focus and classified by what the provider publishes.
URLs last verified 2026-05-26.

\paragraph{Provider clusters.} Six clusters, each capturing a different point of view on the market rather than a clean structural distinction:
\begin{itemize}\setlength{\itemsep}{1pt}
\item \textit{Hyperscalers}: multi-region, multi-service IaaS+PaaS+SaaS portfolios with global reach.
\item \textit{Mid-tier and regional IaaS}: general-purpose IaaS with a narrower geographic footprint or service catalog than the hyperscalers, often positioned on jurisdiction or sovereignty (e.g., EU data residency).
\item \textit{CDN / edge}: globally-distributed edge networks selling caching, request handling, and short-lived compute.
\item \textit{Neocloud}: AI-specialized compute providers with focus on GPUs and other accelerators.
\item \textit{PaaS}: developer-facing application hosting layered atop a hyperscaler. Heroku is excluded as it is being sunset under Salesforce, and broader SaaS / data-platform companies (Salesforce, Snowflake, Databricks, etc.) are out of scope for this paper.
\item \textit{Sustainability-positioned regionals}: regional IaaS or DC operators whose primary differentiator is environmental, e.g. Nordic siting, heat reuse, hydro-powered grids.
\end{itemize}

\paragraph{Status taxonomy.} Five mutually-exclusive disclosure clusters:
\begin{itemize}\setlength{\itemsep}{1pt}
\item \statusA{Per-customer tool}: first-party dashboard or API exposing workload-level emissions per tenant.
\item \statusB{Assured GHG report}: Scope~1/2/3 corporate inventory with external assurance (SBTi-validated, CSRD-baselined, ISO~14064-certified, or otherwise third-party-audited). The corporate number is reliable enough to inherit downstream as Scope~3.
\item \statusB{Self-reported}: published Scope~1/2/3 inventory, \emph{or} substantive quantitative operational disclosure (per-DC carbon intensity, PUE, energy mix, EMAS environmental statement), without external GHG assurance.
\item \statusC{Marketing claim}: forward net-zero pledge, ``100\,\% renewable'' / REC-backed statement, or qualitative sustainability strategy, but no quantified disclosure.
\item \statusC{No disclosure}: nothing surfaced through our search.
\end{itemize}

\begin{longtable}{@{}p{2.5cm}p{1.5cm}p{3.0cm}p{9.5cm}@{}}
\toprule
\textbf{Provider} & \textbf{Region} & \textbf{Status} & \textbf{Reference} \\
\midrule
\endfirsthead
\toprule
\textbf{Provider} & \textbf{Region} & \textbf{Status} & \textbf{Reference} \\
\midrule
\endhead
\bottomrule
\endfoot
\multicolumn{4}{@{}l}{\textit{Hyperscalers}} \\
\midrule
AWS                       & US      & \statusA{Per-customer tool}        & \href{https://aws.amazon.com/sustainability/tools/console/}{aws.amazon.com/sustainability/tools/console/} \\
Alibaba Cloud             & China   & \statusA{Per-customer tool}        & \href{https://alibabacloud.com/en/product/energy-expert}{alibabacloud.com/en/product/energy-expert} \\
Azure                     & US      & \statusA{Per-customer tool}        & \href{https://microsoft.com/en-us/sustainability/emissions-impact-dashboard}{microsoft.com/en-us/sustainability/emissions-impact-dashboard} \\
GCP                       & US      & \statusA{Per-customer tool}        & \href{https://docs.cloud.google.com/carbon-footprint/docs/methodology}{docs.cloud.google.com/carbon-footprint/docs/methodology} \\
IBM Cloud                 & US      & \statusA{Per-customer tool}        & \href{https://ibm.com/cloud/carbon-calculator}{ibm.com/cloud/carbon-calculator} \\
Oracle (OCI)              & US      & \statusA{Per-customer tool}        & \href{https://docs.oracle.com/iaas/Content/General/Concepts/emissions-management.htm}{docs.oracle.com/iaas/Content/General/Concepts/emissions-management.htm} \\
Huawei Cloud              & China   & \statusB{Assured GHG}              & \href{https://huawei.com/en/sustainability}{huawei.com/en/sustainability} \\
Tencent Cloud             & China   & \statusB{Self-reported}             & \href{https://tencentcloud.com/global-infrastructure/sustainability}{tencentcloud.com/global-infrastructure/sustainability} \\
\midrule
\multicolumn{4}{@{}l}{\textit{Mid-tier and regional IaaS}}\\
\midrule
OVHcloud                  & EU-FR   & \statusA{Per-customer tool}        & \href{https://ovhcloud.com/en/about-us/environmental-impact-tracker/}{ovhcloud.com/en/about-us/environmental-impact-tracker/} \\
Scaleway                  & EU-FR   & \statusA{Per-customer tool}        & \href{https://scaleway.com/en/environmental-footprint-calculator/}{scaleway.com/en/environmental-footprint-calculator/} \\
T Cloud Public            & EU-DE   & \statusA{Per-customer tool}        & \href{https://t-cloud-public.com/en/blog/product-news/sustainability-dashboard}{t-cloud-public.com/en/blog/product-news/sustainability-dashboard} \\
Infomaniak                & EU-CH   & \statusB{Assured GHG}              & \href{https://infomaniak.com/en/ecology}{infomaniak.com/en/ecology} \\
Rackspace                 & US      & \statusB{Assured GHG}              & \href{https://www.rackspace.com/resources/2025-sustainability-report}{rackspace.com/resources/2025-sustainability-report} \\
Baidu AI Cloud            & China   & \statusB{Self-reported}             & \href{https://home.baidu.com}{home.baidu.com} \\
Exoscale                  & EU-CH   & \statusB{Self-reported}             & \href{https://exoscale.com/sustainability}{exoscale.com/sustainability} \\
Hetzner                   & EU-DE   & \statusB{Self-reported}             & \href{https://hetzner.com/unternehmen/nachhaltigkeit}{hetzner.com/unternehmen/nachhaltigkeit} \\
JD Cloud                  & China   & \statusB{Self-reported}             & \href{https://ir.jd.com/esgcsr}{ir.jd.com/esgcsr} \\
DigitalOcean              & US      & \statusC{Marketing claim}          & \href{https://digitalocean.com/impact}{digitalocean.com/impact} \\
STACKIT                   & EU-DE   & \statusC{Marketing claim}          & \href{https://stackit.com/en/learn/knowledge/cloud/sustainability}{stackit.com/en/learn/knowledge/cloud/sustainability} \\
Vultr                     & US      & \statusC{No disclosure}            & \na \\
Yandex Cloud              & Russia  & \statusC{No disclosure}            & \na \\
\midrule
\multicolumn{4}{@{}l}{\textit{CDN / edge}} \\
\midrule
Akamai                    & US      & \statusA{Per-customer tool}        & \href{https://akamaisustainability.com/circularity/customer-engagement/}{akamaisustainability.com/circularity/customer-engagement/} \\
Cloudflare                & US      & \statusA{Per-customer tool}        & \href{https://cf-assets.www.cloudflare.com/slt3lc6tev37/7koyyovVxIqK8zdG1pqo6O/dfcbbde950394c61855befd9bf083a28/Impact-Report-2025-Final.pdf}{Cloudflare Impact Report 2025} \\
Fastly                    & US      & \statusA{Per-customer tool}        & \href{https://fastly.com/products/sustainability-dashboard}{fastly.com/products/sustainability-dashboard} \\
Gcore                     & EU-LU   & \statusC{Marketing claim}          & \href{https://gcore.com/about}{gcore.com/about} \\
Bunny.net                 & EU-SI   & \statusC{No disclosure}            & \na \\
CDNetworks                & China   & \statusC{No disclosure}            & \na \\
\midrule
\multicolumn{4}{@{}l}{\textit{Neocloud}} \\
\midrule
Crusoe                    & US      & \statusB{Self-reported}             & \href{https://www.crusoe.ai/resources/blog/crusoe-2024-impact-report}{crusoe.ai/resources/blog/crusoe-2024-impact-report} \\
Nebius                    & EU-NL   & \statusB{Self-reported}             & \href{https://nebius.com/sustainability}{nebius.com/sustainability} \\
Verda                     & Nordic  & \statusC{Marketing claim}        & \href{https://verda.com/}{verda.com} \\
Hyperstack                & UK      & \statusC{Marketing claim}          & \href{https://www.hyperstack.cloud/why-hyperstack}{hyperstack.cloud/why-hyperstack} \\
CoreWeave                 & US      & \statusC{No disclosure}            & \na \\
Lambda                    & US      & \statusC{No disclosure}            & \na \\
Fireworks AI              & US      & \statusC{No disclosure}            & \na \\
Paperspace                & US      & \statusC{No disclosure}            & \na \\
Replicate                 & US      & \statusC{No disclosure}            & \na \\
RunPod                    & US      & \statusC{No disclosure}            & \na \\
Together AI               & US      & \statusC{No disclosure}            & \na \\
\midrule
\multicolumn{4}{@{}l}{\textit{PaaS}} \\
\midrule
Netlify                   & US      & \statusC{Marketing claim}          & \href{https://netlify.com/sustainability}{netlify.com/sustainability} \\
Vercel                    & US      & \statusC{Marketing claim}          & \href{https://vercel.com/kb/guide/what-is-vercel-green-energy-policy}{vercel.com/kb/guide/what-is-vercel-green-energy-policy} \\
Fly.io                    & US      & \statusC{No disclosure}            & \na \\
Railway                   & US      & \statusC{No disclosure}            & \na \\
Render                    & US      & \statusC{No disclosure}            & \na \\
\midrule
\multicolumn{4}{@{}l}{\textit{Sustainability-positioned regionals}} \\
\midrule
atNorth                   & Nordic  & \statusB{Assured GHG}              & \href{https://www.atnorth.com/wp-content/uploads/2025/04/Sustainability-Report-2024-web.pdf}{atnorth.com/.../Sustainability-Report-2024} \\
GleSYS                    & Sweden  & \statusB{Assured GHG}              & \href{https://glesys.com/sustainability}{glesys.com/sustainability} \\
UpCloud                   & Finland & \statusB{Self-reported}             & \href{https://upcloud.com/esg-report-2024/}{upcloud.com/esg-report-2024} \\
Verne                     & Nordic  & \statusB{Self-reported}             & \href{https://verne.co/about-us/sustainability}{verne.co/about-us/sustainability} \\
Leafcloud                 & EU-NL   & \statusC{Marketing claim}          & \href{https://leaf.cloud/truly-green}{leaf.cloud/truly-green} \\
\end{longtable}

\section{Scope-3 Category Coverage in the Big-3 Customer Tools}
\label{app:bigthree-s3}

No two of the three hyperscaler customer tools count the same set of Scope-3 categories (\S\ref{sec:sota}), so even their headline totals are not directly comparable.

\begin{table}[h]
\centering
\caption{Scope-3 category coverage in the Big-3 customer tools, by GHG-Protocol
category. Cat~2 is split into IT hardware vs.\ building embodied. Lifetime amortization
differs: AWS 6\,yr IT~+~50\,yr buildings; GCP \textbf{4\,yr} IT (acknowledged shorter
than actual); Azure 6\,yr IT, buildings excluded. All three zero out past amortization.
\yes\ included, \no\ excluded, \maybe\ partial, \na\ not applicable.}
\label{tab:bigthree-s3}
\footnotesize
\setlength{\tabcolsep}{4pt}
\begin{tabular}{@{}l c c c@{}}
\toprule
GHG-Protocol Scope-3 category & AWS & GCP & Azure \\
\midrule
Cat~1 --- Purchased goods \& services & \na & \na & \yes \\
Cat~2 --- IT hardware (cradle-to-gate) & \yes & \yes & \yes \\
Cat~2 --- Building embodied & \yes & \yes & \no \\
Cat~3 --- FERA (upstream fuel + T\&D losses) & \yes & \yes & \no \\
Cat~4 --- Upstream transport & \yes & \maybe & \yes \\
Cat~5 --- Waste in operations & \no & \no & \yes \\
Cat~6 --- Business travel & \no & \yes & \no \\
Cat~7 --- Employee commuting & \no & \yes & \no \\
Cat~9 --- Downstream transport & \no & \no & \yes \\
Cat~12 --- End-of-life treatment & \no & \no & \yes \\
\bottomrule
\end{tabular}
\end{table}

\end{document}
